---
title: "TidyTuesday - Age Gap in Hollywood"
author: "Kainalytics"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Download Data

```{r}
library(tidytuesdayR)
library(lubridate)
colors_palette <- pal('flat')

tuesdata <- tidytuesdayR::tt_load(2023, week = 7)

age_gaps <- tuesdata$age_gaps
```

# Data Exploration
```{r}
library(tidyverse)
source('appearances.R')
age_gaps %>% 
  glimpse()
```

## movie_name
```{r}
age_gaps %>% 
  count(movie_name, sort = T)


```
For now, I don't know what to do with movie names, except we can see that there are 7 instances of the movie 'Love Actually'. Let's see what we can guess about wht 7 instances it could be
```{r}
age_gaps %>% 
  filter(movie_name == 'Love Actually') %>% 
  glimpse()
```
Alright so there's apparently 7 different couples in this film, hence the 7 occurances. 

## release_year
```{r}
age_gaps %>% 
  count(release_year, sort = T)

age_gaps %>% 
  ggplot(aes(x = release_year))+
  geom_histogram()+
  scale_x_continuous(breaks = c(seq(1920, 2025, by = 5)))
```
Most films data comes from after 1995ish


## director
```{r}
age_gaps %>% 
  count(director, sort = T)

age_gaps %>% 
  mutate(director = fct_lump(director, n = 5)) %>% 
  count(director, sort = T, name = 'No. of films')

age_gaps %>% 
  filter(director == 'Woody Allen') %>% 
  count(movie_name)
```
Woody Allen is the most common director in this dataset

## age_difference
```{r}
age_gaps %>% 
  ggplot(aes(x = age_difference)) +
  geom_histogram()
```
The distribution is very right skewed but does have long tails. 

## couple_number
```{r}
age_gaps %>% 
  count(couple_number, sort = T)

age_gaps %>% 
  ggplot(aes(x = couple_number)) +
  geom_histogram()
  
```
Unsurprisingly most films have a single couple. 

## character_1_gender and character_2_gender
```{r}
age_gaps %>% 
  count(character_1_gender)

age_gaps %>% 
  count(character_2_gender)

age_gaps %>% 
  count(character_1_gender, character_2_gender)
```
12 films have an all male couple, 11 have an all female couple. Unsure as yet as to why man-woman and woman-man are different. 

UPDATE: From the data dictionary, man-woman means the man is older than the woman in this couple and woman-man means the woman is older in the couple. 

## actor_1_birthdate and actor_2_birthdate
```{r}
age_gaps %>% 
  count(yr = year(actor_1_birthdate)) %>% 
  ggplot(aes(x = yr, y = n))+
  geom_col()

age_gaps %>% 
  count(yr = year(actor_1_birthdate), sort = T)

age_gaps %>% 
  count(yr = year(actor_2_birthdate)) %>% 
  ggplot(aes(x = yr, y = n))+
  geom_col()

age_gaps %>% 
  count(yr = year(actor_2_birthdate), sort = T)

age_gaps %>% 
  count(yr = year(actor_1_birthdate)) %>%
  summarise(mean_younger_actor_yr = weighted.mean(yr, n))

age_gaps %>% 
  count(yr = year(actor_2_birthdate)) %>%
  summarise(mean_younger_actor_yr = weighted.mean(yr, n))
```
Most of the older actors between the couple were born in the 1960s and 1970s. 1964 being the most common year with 57 older actor in the couple, which is mostly males.  Weighted mean of the older actor's birth year seems to be 1960. 

Similarly, most of the younger actors among the couple were born in the 1970s. 1976 being the most common year with 48 younger actors being born then, most of them female. Weighted mean of the younger actor's birth year seems to be 1971

# Insights

## Gather
```{r}
yrly <- age_gaps %>% 
  mutate(a1_yr = year(actor_1_birthdate),
         a2_yr = year(actor_2_birthdate)) %>% 
  gather(metric, value, -release_year)

y2 <- yrly %>% 
  filter(metric == 'age_difference') %>% 
  mutate(value = as.numeric(value)) %>% 
  group_by(release_year) %>% 
  na.omit() %>% 
  summarise(mean_age_difference = mean(value)) 

mmn <- mean(y2$mean_age_difference)

y2 %>% 
  ggplot(aes(x = release_year, y = mean_age_difference)) +
  geom_smooth(color = 'black') +
  scale_x_continuous(breaks = c(seq(min(yrly$release_year), 
                                    max(yrly$release_year), by = 5))) +
  expand_limits(y = 0) +
  geom_hline(yintercept =  mmn, color = 'orange', linetype = 'twodash')

  
```

# Men and Women
```{r}
hetero <- age_gaps %>% 
  group_by(movie_name) %>%
  select(release_year, age_difference, couple_number, character_1_gender, 
         character_2_gender, actor_1_age, actor_2_age) %>% 
  mutate(couple_type = paste0(character_1_gender, character_2_gender)) %>% 
  filter(couple_type %in% c('manwoman', 'womanman')) %>% 
  mutate(male_age = case_when(couple_type == 'manwoman' ~ actor_1_age,
                        couple_type == 'womanman' ~ actor_2_age),
         female_age = case_when(couple_type == 'manwoman' ~ actor_2_age,
                        couple_type == 'womanman' ~ actor_1_age)) %>% 
  select(release_year, couple_number, age_difference, male_age, female_age)


hetero %>% 
  ungroup() %>% 
  group_by(release_year) %>% 
  summarise(mean_male_age = mean(male_age),
            mean_female_age = mean(female_age)) %>% 
  mutate(mean_age_difference = mean_male_age - mean_female_age) %>% 
  filter(mean_age_difference < 0)
```
We see that in 1971 mean female age was higher than the mean male age, so let's analyze films in 1971
 
# Films in 1971
```{r}
age_gaps %>% 
  filter(release_year == 1971) %>% 
  glimpse()
```
Okay, so it's being skewed by only 1 out of only 3 observations for the film 'Harold and Maude'

## Harold and Maude
```{r}
age_gaps %>% 
  filter(movie_name == 'Harold and Maude') %>% 
  glimpse()
```
That was interesting, now back to the issue at hand

## Male Female age difference
```{r}
hetero_long <- hetero %>% 
  ungroup() %>% 
  group_by(release_year) %>% 
  summarise(mean_male_age = mean(male_age),
            mean_female_age = mean(female_age)) %>% 
  mutate(mean_age_difference = mean_male_age - mean_female_age) %>% 
  gather(metric, value, -c(release_year, mean_age_difference)) %>% 
  arrange(release_year)


maxage <- hetero_long %>% 
  filter(release_year == min(release_year), 
         row_number() == 1) %>% 
  pull(mean_age_difference)
minage <- hetero_long %>% 
  filter(release_year == max(release_year)) %>% 
  filter(row_number() == 1) %>% 
  pull(mean_age_difference)

  
maxyr <- max(hetero_long$release_year)
minyr <- min(hetero_long$release_year)


hetero_long %>% 
  ggplot()+
  geom_col(mapping = aes(x = release_year, y = value, fill = metric), 
           position = 'fill') +
  scale_fill_manual(values=c(colors_palette[6], colors_palette[3], 
                             "#899999",  "#96B4E9", "#E19F00"), 
                       name="Share in Average Age of",
                       breaks=c("mean_female_age", "mean_male_age"),
                       labels=c("Female Actors", "Male Actors")) +
  geom_smooth(mapping = aes(x = release_year, y =0.05 * mean_age_difference), 
              color = 'black') +
  annotate(x = minyr, y = maxage * 0.05 - 0.45, label = paste('Age Gap', maxage, 'yrs', sep = " "), geom = 'text') +
  annotate(x = maxyr, y = minage * 0.05 + 0.05, label = paste('Age Gap', minage, 'yrs', sep = " "), geom = 'text') +
  theme(legend.position = 'bottom', legend.background = element_blank(),
        legend.key.height = unit(.1,'line'), legend.title = element_text(size = 10),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        panel.border = element_rect(colour = colors_palette[9], fill=NA, size=1)) +
  labs(x = '',
       y = '',
       title = 'Narrowing Gap between the ages of male and female actors portraying a couple in Hollywood films',
       subtitle = 'Heterosexual Couples Only | Data until 2022',
       caption = 'The data this week comes from Hollywood Age Gap via Data Is Plural.\n Analysis: @sachasucha\n
       Note: The age gaps dataset includes "gender" columns, which always contain the values "man" or "woman".These values appear to indicate how the characters in each film identify. Some of these values do not match how the actor identifies.') +
  scale_x_continuous(breaks = c(seq(minyr, maxyr, by = 3)))
  
```

# Alt text for Twitter
Plot title: "Narrowing Gap between the ages of male and female actors portraying a couple in Hollywood films"
Subtitle: Heterosexual Couples Only | Data until 2022
Tag: #TidyTuesday | Data: Hollywood Age Gap Via Data Is Plural


Filled Stack Plot showing Average Age of Female actors in blue, and that of male actors in yellow. A smoothed line is overlaid showing the age gap reducing from 28.5 years average in 1930s to under 5 years in 2022.

Plot is themed in a flat style palette. Plot background is a light grey whereas all text is in black and the stacked bar plots are yellow and blue. 
